A boundary-integral framework to simulate \\ the viscous erosion of a porous medium

### 2018-04-06 docs/methods/erosion_methods.tex
added paragraph headings

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
To perform this simulation of Stokesian erosion of a porous media, two
fast and accurate methods are employed---one to solve the fluid
equations and the other to integrate the eroding grains.  The fluid
equations are reformulated in terms of a boundary integral equation
(BIE)~\cite{poz1992}.  Since the boundary conditions are Dirichlet, we
use a double-layer potential formulation so that the resulting BIE is
second-kind.  This second-kind BIE is discretized with the spectrally
accurate trapezoid rule~\cite{tre-wei2014} which results in a
well-conditioned dense linear system for an unknown density function.
This linear system is solved using fast-multipole-accelerated
GMRES~\cite{saa-sch1986}.  Once the density function is computed, the
erosion rate (shear stress) is computed with a spectrally accurate
odd-even quadrature rule~\cite{sid-isr1988}.  In addition, the pressure
and vorticity are computed in the fluid bulk (Fig.~\ref{fig:50bodies})
with an upsampled trapezoid rule in the fluid bulk to assist with
visualizing the effect of erosion in porous media.

@@

The effect of erosion on a geometry is quite rich.  For instance, it has
been hypothesized and numerically observed that the erosion leads to a
shear stress that is uniform on the body which results in a uniform
erosion rate.  The shape that has a uniform shear stress is analyzed in
three dimensions by Montenegro-Johnson and Lauga~\cite{mon-lau2015}
which is an extension of the analysis done by Pironneau~\cite{pir1973}
to find a Stokesian drag minimizing shape. 
%\todo[inline]{Double check this statement. I am not sure that the shape found by Lauga has uniform shear stress. The shape found by Pironneau does.}
In this idealized case, analytical axisymmetric shapes that minimize the
drag can be found.  However, the effect of erosion on a porous media is
less understood.  The work most closely related to ours simulates erosion in a three-dimensional incompressible Stokesian fluid~\cite{mit-spa2016}, but they only consider single eroding body.

@@

We make two fundamental design choices: first, we restrict ourselves to
two spatial dimensions, and second the bodies are considered immobile
(i.e.~fixed to their initial locations). The restriction to two
dimensions allows us to simulate the erosion of many bodies (up to $50$
in this manuscript) with high resolution, giving us the ability to
accurately characterize the erosion of a porous medium. While BIEs are
available in three dimensions, the extra computational cost limits the
number of bodies that could be practically simulated, making
porous-medium simulations less feasible. The second assumption of
immobile bodies affords a simple framework to study the effects of
erosion in isolation. While other effects, such as sedimentation or
transport, undoubtedly play a role in real applications, these effects
would introduce additional complexity, both conceptually and
numerically. We therefore defer the study of these effects to future
work.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In Section~\ref{s:formulation}, we describe the governing equations for
both the fluid and the erosion model in a \thL~framework.  In
Section~\ref{s:method}, the numerical methods in both space and time are
described and Section~\ref{s:qoi} describes how to compute the
vorticity, pressure, and drag.  Section~\ref{s:SingleResults} gives
several numerical examples for a single body, and
Section~\ref{s:MultiResults} presents numerical examples for
multiply-connected domains.  Finally, concluding remarks are made in
Section~\ref{s:conclusions}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Contributions}
To perform this simulation of Stokesian erosion of a porous media, two
fast and accurate methods are employed---one to solve the fluid
equations and the other to integrate the eroding grains.  The fluid
equations are reformulated in terms of a boundary integral equation
(BIE)~\cite{poz1992}.  Since the boundary conditions are Dirichlet, we
use a double-layer potential formulation so that the resulting BIE is
second-kind.  This second-kind BIE is discretized with the spectrally
accurate trapezoid rule~\cite{tre-wei2014} which results in a
well-conditioned dense linear system for an unknown density function.
This linear system is solved using fast-multipole-accelerated
GMRES~\cite{saa-sch1986}.  Once the density function is computed, the
erosion rate (shear stress) is computed with a spectrally accurate
odd-even quadrature rule~\cite{sid-isr1988}.  In addition, the pressure
and vorticity are computed in the fluid bulk (Fig.~\ref{fig:50bodies})
with an upsampled trapezoid rule in the fluid bulk to assist with
visualizing the effect of erosion in porous media.

@@

\paragraph{Related Work}
The effect of erosion on a geometry is quite rich.  For instance, it has
been hypothesized and numerically observed that the erosion leads to a
shear stress that is uniform on the body which results in a uniform
erosion rate.  The shape that has a uniform shear stress is analyzed in
three dimensions by Montenegro-Johnson and Lauga~\cite{mon-lau2015}
which is an extension of the analysis done by Pironneau~\cite{pir1973}
to find a Stokesian drag minimizing shape. 
%\todo[inline]{Double check this statement. I am not sure that the shape found by Lauga has uniform shear stress. The shape found by Pironneau does.}
In this idealized case, analytical axisymmetric shapes that minimize the
drag can be found.  However, the effect of erosion on a porous media is
less understood.  The work most closely related to ours simulates erosion in a three-dimensional incompressible Stokesian fluid~\cite{mit-spa2016}, but they only consider single eroding body.

@@

\paragraph{Limitations}
We make two fundamental design choices: first, we restrict ourselves to
two spatial dimensions, and second the bodies are considered immobile
(i.e.~fixed to their initial locations). The restriction to two
dimensions allows us to simulate the erosion of many bodies (up to $50$
in this manuscript) with high resolution, giving us the ability to
accurately characterize the erosion of a porous medium. While BIEs are
available in three dimensions, the extra computational cost limits the
number of bodies that could be practically simulated, making
porous-medium simulations less feasible. The second assumption of
immobile bodies affords a simple framework to study the effects of
erosion in isolation. While other effects, such as sedimentation or
transport, undoubtedly play a role in real applications, these effects
would introduce additional complexity, both conceptually and
numerically. We therefore defer the study of these effects to future
work.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Outline of the Paper}
In Section~\ref{s:formulation}, we describe the governing equations for
both the fluid and the erosion model in a \thL~framework.  In
Section~\ref{s:method}, the numerical methods in both space and time are
described and Section~\ref{s:qoi} describes how to compute the
vorticity, pressure, and drag.  Section~\ref{s:SingleResults} gives
several numerical examples for a single body, and
Section~\ref{s:MultiResults} presents numerical examples for
multiply-connected domains.  Finally, concluding remarks are made in
Section~\ref{s:conclusions}.
